Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-05 ms.tex
fixed abstract2

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average star formation history of galaxies containing a given
black hole mass.  Again, Ia heating is suppressed in low mass galaxies
because the Ia radius exceeds the nominal stagnation radius (bottom
panel).  MSP heating is again found to be negligible for all black
hole masses.  The younger stellar populations characterizing galaxies
with $\Mbh\lsim 5\times 10^{8} \Msun$ are dominated by stellar wind
and Ia heating, with $v_{w}^{\star} \sim 1000$ km s$^{-1}$ and
$v_w^{\rm Ia} \sim 500$ km s$^{-1}$, respectively.  Only for the most
massive galaxies, with $\Mbh\gsim 5\times 10^{8} \Msun$, does the lack
of young stellar populations significantly reduce the role of stellar
wind
heating.  %Solid and dashed lines again show black hole masses for which the flow is thermally stable and thermally unstable, respectively, with $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ as a function of $M_{\bullet}$ shown in the bottom panel.
Strikingly, {\it we find that galactic nuclei experiencing the same
  average star formation history as their host galaxies produce
  thermally stable flows across almost all black hole masses}.  One exception may occur for the highest mass black holes, for which the total heating velocity is comparable to the stellar velocity dispersion ($\zeta \sim \zeta_c$) and thermal stability for the average star formation history is marginal, especially for core galaxies ($\S\ref{sec:mdot}$).

@@

Figure \ref{fig:bh_growth} (top panel) shows with solid lines the
accretion rate as a function of black hole mass for the average star
formation heating, calculated from equation (\ref{eq:eddr_analytic})
using our results for the total wind heating
(Fig.~\ref{fig:mdot_mass}).  This average accretion rate increases
from $\dot{M} \sim 10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ for low-mass
black holes ($M_{\bullet} \sim 10^{6}M_{\odot}$) to $\dot{M} \sim
10^{-4}M_{\odot}$ yr$^{-1}$ for $M_{\bullet} \sim 10^{9}M_{\sun}$.
These fall below the maximum thermally stable accretion rate,
$\dot{M}_{\rm TI}$ (yellow lines), except at the highest
black hole masses $\sim 10^{9}M_{\odot}$.  For these highest mass
black holes, the maximum accretion rate as set by the SN Ia
radius (green lines; eq.~[\ref{eq:eddr_Ia}]) could instead cap the
accretion rate close to the thermally-unstable value (see discussion following eq.~[\ref{eq:eddr_Ia}]).  For lower mass black holes, the maximum accretion rate set by SNe Ia is significantly higher and cannot
also prevent the development of thermally unstable flows,
i.e. $\dot{M}_{\rm Ia} \gg \dot{M}_{\rm TI}$.

Shown also for comparison in Figure \ref{fig:bh_growth} (top panel) is the accretion rate onto SgrA* calculated by \citet{Quataert:2004a}, as well as the range of $\dot{M}$ for the low-luminosity AGN NGC3115 from \citealt{ShcherbakovWong+:2014a} (who find a range of accretion rates depending on the assumed model for thermal conductivity, although they do not include saturation of the conductive flux).  The SgrA* rate exceeds our estimate of $\dot{M}$ due to the average star formation history for galaxies with the same mass black hole by almost two orders of magnitude.  This discrepancy can be understood by the fact that star formation in the central parsec of the Galaxy is not in steady-state: feedback is instead dominated by the stellar winds from the ring of young massive stars of mass $\sim 10^{4}M_{\odot}$ and estimated age $\sim 10$ Myr (e.g., \citealt{Schodel+07}).  Such a star formation history is better described as a `burst' of stars, for which the value of $\eta \sim 100$ at $\tau_{\star} = 10^{7}$ yr (Fig.~\ref{fig:vwImp}) is a factor of $\sim 100$ times higher than the value of $\eta \sim 1$ predicted in the average SFH case (Fig.~\ref{fig:vwSources}, bottom panel).

@@

The (unresolved) core X-ray luminosities of galactic nuclei provide a
powerful diagnostic of the SMBH accretion rates and how they vary with
SMBH mass and other galaxy properties (e.g., \citealt{Ho08} and
references therein).  Figure~\ref{fig:miller} shows our predictions
for $\langle L_{X} \rangle =\epsilon_X \dot{M} c^2$ as a function of
black hole mass, where $\epsilon_{X} = \epsilon\epsilon_{\rm bol}$,
$\epsilon$ is the total radiative efficiency, and $\epsilon_{\rm bol}
\sim 0.1$ is the bolometric correction into the measured X-ray band
for low luminosity AGN (\citealt{Ho08}).  The value of $\langle L_X
\rangle$ is calculated using the average accretion for galaxies with
core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) stellar density
profiles, using the relationship $\Gamma = 0.3 \Mbheight^{-0.24}$ fit
to the \citet{LauerFaber+:2007a} galaxy sample.

>>>>

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average star formation history of galaxies containing a given
black hole mass.  Again, Ia heating is suppressed in low mass galaxies
because the Ia radius exceeds the nominal stagnation radius (bottom
panel).  MSP heating is again found to be negligible for all black
hole masses.  The younger stellar populations characterizing galaxies
with $\Mbh\lsim 5\times 10^{8} \Msun$ are dominated by stellar wind
and Ia heating, with $v_{w}^{\star} \sim 1000$ km s$^{-1}$ and
$v_w^{\rm Ia} \sim 500$ km s$^{-1}$, respectively.  Only for the most
massive galaxies, with $\Mbh\gsim 5\times 10^{8} \Msun$, does the lack
of young stellar populations significantly reduce the role of stellar
wind
heating.  %Solid and dashed lines again show black hole masses for which the flow is thermally stable and thermally unstable, respectively, with $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ as a function of $M_{\bullet}$ shown in the bottom panel.
Strikingly, {\it we find that galactic nuclei experiencing the same
  average star formation history as their host galaxies produce
  thermally stable flows across almost all black hole masses}.  One important exception occurs for the highest mass black holes, for which the total effective wind velocity is comparable to the stellar velocity dispersion ($\zeta \sim \zeta_c$) and thermal stability for the average star formation history is marginal, especially for core galaxies ($\S\ref{sec:mdot}$).

@@

Figure \ref{fig:bh_growth} (top panel) shows with solid lines the
accretion rate as a function of black hole mass for the average star
formation heating, calculated from equation (\ref{eq:eddr_analytic})
using our results for the total wind heating
(Fig.~\ref{fig:mdot_mass}).  This average accretion rate increases
from $\dot{M} \sim 10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ for low-mass
black holes ($M_{\bullet} \sim 10^{6}M_{\odot}$) to $\dot{M} \sim
10^{-4}M_{\odot}$ yr$^{-1}$ for $M_{\bullet} \sim 10^{9}M_{\sun}$.
These fall below the maximum thermally stable accretion rate,
$\dot{M}_{\rm TI}$ (yellow lines), except at the highest
black hole masses $\sim 10^{9}M_{\odot}$.

For low mass black holes, blow out from near the SN Ia radius caps the SMBH accretion rate at a value $\dot{M}_{\rm Ia} \sim 10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ (green lines; eq.~[\ref{eq:eddr_Ia}]), which is however not low enough to prevent otherwise thermally-unstable flow at smaller radii (see discussion following eq.~[\ref{eq:eddr_Ia}]).  For higher mass black holes the stagnation resides close to the Ia radius, but the energy released by a single Ia is insufficient to unbind gas from the stellar bulge (although Ia heating contributes to the steady heating rate).

%For these highest mass
%black holes, the maximum accretion rate as set by the SN Ia
%radius (green lines; eq.~[\ref{eq:eddr_Ia}]) could instead cap the
%accretion rate close to the thermally-unstable value (see discussion following eq.~[\ref{eq:eddr_Ia}]).  For lower mass black holes, the maximum accretion rate set by SNe Ia is significantly higher and cannot
%also prevent the development of thermally unstable flows,
%i.e. $\dot{M}_{\rm Ia} \gg \dot{M}_{\rm TI}$.

Shown also for comparison in Figure \ref{fig:bh_growth} (top panel) is the accretion rate onto SgrA* calculated by \citet{Quataert:2004a}, as well as the range of $\dot{M}$ for the low-luminosity AGN NGC3115 derived through detailed modeling by \citealt{ShcherbakovWong+:2014a} (who find a range of accretion rates depending on the assumed model for thermal conductivity, although they do not include saturation of the conductive flux as considered in this work).  The SgrA* rate exceeds our estimate of $\dot{M}$ due to the average star formation history for the same $M_{\bullet}$ by almost two orders of magnitude.  This discrepancy can be understood by the fact that star formation in the central parsec of the Milky Way is not in steady-state: feedback is instead dominated by the stellar winds from the ring of young massive stars of mass $\sim 10^{4}M_{\odot}$ and estimated age $\sim 10$ Myr (e.g., \citealt{Schodel+07}).  Such a star formation history is better described by our `star burst' scenario, for which the value of $\eta \sim 100$ at $\tau_{\star} = 10^{7}$ yr (Fig.~\ref{fig:vwImp}) is a factor of $\sim 100$ times higher than the value of $\eta \sim 1$ predicted in the average SFH case (Fig.~\ref{fig:vwSources}, bottom panel).

@@

The unresolved core X-ray luminosities of nearby galactic nuclei provide a
powerful diagnostic of the SMBH accretion rates and how they vary with
SMBH mass and other galaxy properties (e.g., \citealt{Ho08} and
references therein).  Figure~\ref{fig:miller} shows our predictions
for $\langle L_{X} \rangle =\epsilon_X \dot{M} c^2$ as a function of
black hole mass, where $\epsilon_{X} = \epsilon\epsilon_{\rm bol}$,
$\epsilon$ is the total radiative efficiency, and $\epsilon_{\rm bol}
\sim 0.1$ is the bolometric correction into the measured X-ray band
for low luminosity AGN (\citealt{Ho08}).  The value of $\langle L_X
\rangle$ is calculated using the average accretion for galaxies with
core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) stellar density
profiles, using the relationship $\Gamma = 0.3 \Mbheight^{-0.24}$ fit
to the \citet{LauerFaber+:2007a} galaxy sample.
